Last glacial cycle glacier erosion potential in the Alps

### 2021-05-26 articles/erosion/response-referee2.md

    Lateral stress gradients missing from the shallow-shelf approximation
    stress balance could also contribute to moderate sliding velocity in narrow
    troughs (Herman et al., 2011; Egholm et al., 2012a, b; Pedersen et al.,
    2014).

> It might be pointed out that the ‘Koppes’ relationship with an exponent of
> 2.34 is based on 13 data points that are very poorly distributed: in fact,
> Koppes et al. (2015) go on to drop two outliers and reach an exponent of
> 2.62.

We clarified in the methods that we use  Koppes et al. (2015) erosion law
deriving from their full dataset. In addition, this sentence was added in the
discussion of the "choice of erosion law":

    On the other extreme, an even more non-linear erosion law, not tested here,
    derived from tidewater glaciers but excluding two outliers (ė = 5.3×10^-9
    u_b^2.62, Koppes et al., 2015) would result in an even more localized
    pattern of erosion potential.

@@

> Fig. 1c illustrates a very unusual cirque, with a huge lake replacing a very
> unhealthy convex glacier. Fascinating, but surely a more representative
> cirque should be used. Actually these three photos are not well integrated
> with the text (pace lines 191 & 195-198). Lauterbrunnen is an extreme
> (vertical-sided) trough example, over-used by textbooks.

TODO: think about replacing the photos.

> Fig. 5: The elevation histogram in Fig. 5a could be misleading, as it seems
> to cover the whole study area, including parts never glacier-covered. It
> should be limited to areas glacier-covered at the maximum, and preferably
> supplemented by a histogram for those covered now or at the minimum. The
> accompanying main graph (5a) relates to areas glacier-covered at each time
> slice. Re another reviewer’s comment, yes LGM glaciers did reach below 500 m:
> at L. Garda (67 m?) and the Tagliamento Glacier lobe LGM ice reached below
> 100 m. An integrated plot of total-cycle potential erosion rate (and
> volume?) at each altitude, over glacier-covered areas, would be interesting
> (as an extra Figure): can rates at high altitude, where glacial duration is
> longer, catch up with those at low altitude?  Probably not, given the
> concentration of flow down major valleys. Striping above 3000 m, presumably
> related to low surface area, suggests a switch there to 20 m or 40 m bands
> would be advisable. Caption; ‘modelled potential erosion rates’?

The figure now includes a histogram of model domain elevations, a histogram of
glacier-covered elevations, and a plot of cumulative erosion potential volume
per elevation bands of 100 m. The latter shows that, as you suspected, high
altitude rapid erosion is largely offset by the limited number of grid cells
it concerns. Instead, the bulk of the erosion potential occurs below 2000 m.

To highlight periods when low-elevation is relevant (and address the comment
from reviewer #1), we also hatched elevation bands that contain fewer than a
hundred ice-covered grid cells for a given time. We also shifted to 100-m
elevation bands on the main panel. Wihle some detail is lost, this avoids
much of the striping (especially on the newly added contour and hatched
pattern), and we find the new plot more readable.

### 2021-05-26 articles/erosion/alps-erosion.tex
Update hypsogram caption and responses.

    \begin{figure*}
      \centerline{\includegraphics{alpero_hypsogram}}
      \caption{%
        \textbf{(a)} Potential erosion rate ``hypsogram'', showing the
          geometric mean of modelled rates in 10-m elevation bands
          across the entire model domain and its time evolution. Hatches
          indicate elevation bands with fewer than a hundred grid cells
          (100\,$km^2$).
        \textbf{(b)} Distribution of model domain bedrock topography (grey
          bars) and glaciated topography (blue bars) in 100-m elevation bands,
          and the corresponding cumulative potential erosion volume (brown
          line).
        \textbf{(c)} Same as Fig.~\ref{fig:cumulative}b.
        \todo{fix labels, test hatching below a minimum number of grid cells}}
      \label{fig:hypsogram}
    \end{figure*}

>>>>

    \begin{figure*}
      \centerline{\includegraphics{alpero_hypsogram}}
      \caption{%
        \textbf{(a)} Potential erosion rate ``hypsogram'', showing the
          geometric mean of modelled rates in 100-m elevation bands
          across the entire model domain and its time evolution. Hatches
          indicate elevation bands with fewer than a hundred grid cells
          (100\,$km^2$).
        \textbf{(b)} Distribution of model domain bedrock topography (grey
          bars) and glaciated topography (blue bars) in 100-m elevation bands,
          and the corresponding cumulative potential erosion volume (brown
          line).
        \textbf{(c)} Same as Fig.~\ref{fig:cumulative}b.
        \todo{fix labels, test hatching below a minimum number of grid cells}}
      \label{fig:hypsogram}
    \end{figure*}
